---
title: "Pricing the Down and Out Call Option (DOC) in Black-Scholes-Merton Model and Heston Model"
author: "Yaxin LI"
date: "2019.02.09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##2. Black-Scholes-Merton Model
###2.2 Pricing DOC
####2.2.1 Analytical Method
```{r}
library(tidyverse)
S     <- 50         # Initial stock value
sigma <- 0.2        # Volatility
r     <- 0.02       # Risk-free rate
t     <- 1          # Maturity
K     <-40          #strike
B     <-35          #barrier

#Get the fomula for vanilla call option
call_BS <- function(S, K, r, sigma, t) {
  d1 <- (log(S/K) + (r + sigma^2 / 2) * t) / sigma / sqrt(t)
  d2 <- d1 - sigma * sqrt(t)
  return(S * pnorm(d1) - K * exp(-r*t) * pnorm(d2))
}

#Get the Pricing fomula for DOC
DOC_price_BS <-function(S, K, B, r, sigma, t){
  return(max(call_BS(S, K, r, sigma, t)-(S/B)^(1-2*r/(sigma^2))*call_BS(B^2/S, K, r, sigma, t),0))
}

#Get the theoretical value for DOC
theoretical=DOC_price_BS(S, K, B, r, sigma, t)

```


####2.2.2 Monte-Carlo Method
#####3. Simulation
```{r}
nb_sim    <- 10^4      # Number of Monte-Carlo simulations
nb_points <- 10^2      # Number of discretization points
payoff    <- c()       # Empty vector of simulated payoffs

#Simulation
for(j in 1:nb_sim){
  x<-c(0)
  x <- S         # Initializing the process
  if (x<= B){
  tem=0
  break
  }
  for(k in 2:nb_points){
    x[k] <- x[k-1] * (1 + r*t/nb_points + sigma*sqrt(t/nb_points) *rnorm(1)) 
    if (x[k] <=  B){
    tem=0
    break
    }
    tem=x[k]   
  }

  payoff[j] <- max(0,tem-K)
}

#Get the MC value for DOC
MC=exp(-r*t)*mean(payoff)
```


#####4. Visualize the simulation process
```{r}
library(MASS)
nb_sim <- 10        # Number of Monte-Carlo simulations
nb_points <- 10^2   # Number of discretization points
S <- 50             # Initial stock value
sigma <- 0.2        # Volatility
r <- 0.02           # Risk-free rate
t <- 1              # Maturity
payoff <- c()       # Empty vector of simulated payoffs
K <-40              #strike
B <-35              #strike

#Generate 10 sample paths
x<-rep(0,(nb_points*nb_sim))
dim(x)<-c(nb_points,nb_sim)

for(j in 1:nb_sim){
  x[1,j]<- S         # Initializing the process
  
  for(k in 2:nb_points){
    x[k,j] <- x[k-1,j] * (1 + r*t/nb_points + sigma*sqrt(t/nb_points) *rnorm(1)) 
  }
}


#plot sample paths
label="B=35"
sim <- data.frame(cbind(1:nb_points, x))
colnames(sim) <- c("Iteration", 1:nb_sim)
sim <- gather(sim, key = Simulation, value = Value, -Iteration)
sim$Simulation <- as.factor(sim$Simulation)
sim %>% ggplot(aes(x = Iteration, y = Value, color = Simulation)) + 
  geom_line() + theme(text = element_text(size = 13))+geom_hline(aes(yintercept=35),color ="red",linetype="dashed",size=1)+geom_hline(aes(yintercept=40),linetype="dashed",size=1)


```



####2.2.3 Monte-Carlo Method with Refined Barriers
```{r}
payoff <- c()       # Empty vector of simulated payoffs

#Simulation
for(j in 1:nb_sim){
  x<-c(0)
  x <- S         # Initializing the process
  if (x <=  B*exp(-0.5826*sigma*sqrt(1/nb_points))){
  tem=0
  break
  }  
  for(k in 2:nb_points){
    x[k] <- x[k-1] * (1 + r*t/nb_points + sigma*sqrt(t/nb_points) *rnorm(1)) #类似几何收益率的做法.
    if (x[k] <=  B*exp(-0.5826*sigma*sqrt(1/nb_points))){
    tem=0
    break
    }
    tem=x[k]   
  }

  
  payoff[j] <- max(0,tem-K)
}

#Get the refined MC value for DOC
MC_r=exp(-r*t)*mean(payoff)

```


####2.2.4 Fourier Transform and Laplace Transform Method
```{r}
#Wiener-Hopf Fourier Inversion
if(!require(invLT)){install.packages("invLT")}
library(invLT)  # Package for the Laplace inversion
k <- log(K/S)
b <- log(B/S)
d <- r - sigma^2/2
alpha <- 0.01   # Exponential tempering
beta <- 0.01    # Exponential tempering


#Definitions of Functions
phi_p <- function(x, a, sigma, q){ # Wiener-Hopf factor phy_plus
  i <- complex(real = 0, imag = 1)
  l1 <- a/sigma^2 + sqrt(a^2+2*sigma^2*q)/sigma^2
  l1/(l1+i*x)
}

phi_m <- function(x, a, sigma, q){ # Wiener-Hopf factor  phy_minus
  i <- complex(real = 0, imag = 1)
  l2 <- (-a/sigma^2) + sqrt(a^2+2*sigma^2*q)/sigma^2
  l2/(l2-i*x)
}

integrand <- function(u, v, a, sigma, q, k, b, alpha, beta){ # CF of option price
  i <- complex(real = 0, imag = 1)
  num <- (exp(-i*u*k - i*v*b) * phi_p(u+v+i*(beta-alpha-1), a, sigma, q) * phi_m(u-i*(alpha+1), a, sigma, q))
  den <-  (i*u + alpha) * (i*v - beta) * (i*u + alpha + 1) * q
  return(num/den)
}

WH <- function(q, a, sigma, k, b, alpha, beta){ #laplace transform
  nu <- 0.01                                             # Hardcoding the parameters for simplicity
  v <- seq(-10, 10, by = nu)                          # Grid
  u <- seq(-10, 10, by = nu)
  z <- sapply(u, integrand, v = v, a = a, sigma = sigma, q = q, k = k, b = b, alpha = alpha, beta = beta) 
  # The columns pertain to u, the rows to v
  z <- Re(z[is.finite(z)])                                                    
  return(1 / (2*pi)^2 * sum( z, na.rm = T) * nu^2)      # First inversion: double Fourier!  
}

test <- function(x) sapply(x, function(y) {WH(y, d, sigma, k, b, alpha, beta)})


#Get the FT value for DOC
FT=S * exp(-alpha*k+beta*b-r*t) * iv.BrC(test, t = 1, nterms = 19, gamma = 0.02)

```


###2.2.5 Robust Analysis for the price of DOC
####2.2.5.1 Summary of the above four pricing methods
```{r}
#OUTPUT
print(paste("Analytical Method:",theoretical))
print(paste("Monte-Carlo Method:",MC))
print(paste("Monte-Carlo Method with Refined Barriers:",MC_r))
print(paste("Fourier Transform and Laplace Transform Method:",FT))

```


####2.2.5.2 Robust Analysis for the price of DOC
Here, I change the parameters so as to see how the price of DOC evolves with price of the underlying stock, the barrier, the volatility, and the risk free rate respectively.
```{r}
#Price variation with initial stock prices
S_test=20:70
price=sapply(S_test,DOC_price_BS,K=K, B=B, r=r, sigma=sigma, t=t)
price_test <- data.frame(S_test, price, "Price")
ggplot(price_test, aes(x = S_test, y = price)) + geom_line(color='red')

#Price variation with Barrier
barrier_test=30+(1:50)/5
price=sapply(barrier_test,DOC_price_BS,S=S,K=K, r=r,sigma=sigma, t=t)
price_test <- data.frame(barrier_test, price, "Price")
ggplot(price_test, aes(x =barrier_test, y = price)) + geom_line(color='red')

#Price variation with volatility
sigma_test=0.1+(1:50)/100
price=sapply(sigma_test,DOC_price_BS,S=S,K=K, B=B, r=r, t=t)
price_test <- data.frame(sigma_test, price, "Price")
ggplot(price_test, aes(x =sigma_test, y = price)) + geom_line(color='red')

#Price variation with risk free rate
r_test=(1:50)/1000
price=sapply(r_test,DOC_price_BS,S=S,K=K, B=B,sigma=sigma, t=t)
price_test <- data.frame(r_test, price, "Price")
ggplot(price_test, aes(x =r_test, y = price)) + geom_line(color='red')


```


##2.3 Greeks
###2.3.1 Computing Greeks
Here, I compute all the Greeks: Delta, Gamma, Vega, Theta, Pho for the special case. Since the closed form of Greeks for DOC are very complicated, here, I use numerical approximation method to compute the Greeks.
The main idea of this numerical method is that, assuming smoothness and using the center differentiation.

```{r}

h <- 0.0001#simulate delta

#Calculate all the Greeks
Delta=(DOC_price_BS(S+h, K, B, r, sigma, t)-DOC_price_BS(S-h, K, B, r, sigma, t))/2/h
Gamma=(DOC_price_BS(S+h, K, B, r, sigma, t)+DOC_price_BS(S-h, K, B, r, sigma, t)-2*DOC_price_BS(S, K, B, r, sigma, t))/2/(h^2)
Vega=(DOC_price_BS(S, K, B, r, sigma+h, t)-DOC_price_BS(S, K, B, r, sigma-h, t))/2/h
Theta=(DOC_price_BS(S, K, B, r, sigma, t)-DOC_price_BS(S, K, B, r, sigma, t-h))/h
Pho=mean(DOC_price_BS(S, K, B, r+h, sigma, t)-DOC_price_BS(S, K, B, r-h, sigma, t))/2/h

#Output
print(paste("Delta: ",Delta))
print(paste("Gamma:",Gamma))
print(paste("Vega:",Vega))
print(paste("Theta:",Theta))
print(paste("Pho:",Pho))

```

###2.3.2 Robust Analysis for Delta
To carry on the robust analysis for Delta, here I let the initial stock price vary from 20 to 70. In addition, I use four time to maturities, which is 3 months, 6 months, 9 months and 1 year to draw the following 4 curves respectively, so as to see the impact of time to maturity on Delta as well.

```{r}
#Test Delta variation with initial stock prices
Delta_price <- function(S,t){
h <- 0.0001#simulate delta
  return((DOC_price_BS(S+h, K, B, r, sigma, t)-DOC_price_BS(S-h, K, B, r, sigma, t))/2/h)
}

#Generate 4 curves for 4 time to maturity
t_test=c(0.25,0.6,0.75,1)
Delta_test1=sapply(S_test,Delta_price,t=t_test[1])
Delta_test2=sapply(S_test,Delta_price,t=t_test[2])
Delta_test3=sapply(S_test,Delta_price,t=t_test[3])
Delta_test4=sapply(S_test,Delta_price,t=t_test[4])
Delta_test=cbind(Delta_test1,Delta_test2,Delta_test3,Delta_test4)

#Plot
test_greek <- data.frame(S_test, Delta_test)
colnames(test_greek) <- c("Stock_price", (1:4)/4)
test_greek<-gather(test_greek, key =Time_to_maturity, value = Value, -Stock_price)
test_greek$Time_to_maturity<- as.factor(test_greek$Time_to_maturity)
test_greek %>% ggplot(aes(x = Stock_price, y = Value, color = Time_to_maturity)) + 
  geom_line() + theme(text = element_text(size = 13))


```


#3. Heston Model
##3.1 Model Presentation
In 1993, considering the phenomenon of volatility clustering, Heston introduced a model in which the variance is modelled by a Cox-Ingersoll-Ross (CIR) process (Heston, S. ,1993), and the dynamics of stochastic volatility and stock price can be write as follows: 
$$dv_t=(a-bv_t)dt+c\sqrt{v_t}dZ_t^1$$
The diffusion is then simply written as 
$$\frac{dS_t}{S_t}=mdt + \sqrt{v_t}dZ_t^2$$
Where $Z_t^1$ and $Z_t^2$ are two standard Q Brownian motions with a correlation coefficient ρ; a, b, c and m have to be positive to make sure that the model is well defined. 


##3.2 Pricing DOC
###3.2.1 Monte-Carlo Method
#####3. Simulation
```{r}
#Simulation
library(MASS)
S <- 50             # Initial stock value
sigma <- 0.2        # Volatility
r <- 0.02           # Risk-free rate
t <- 1              # Maturity
K <-40 #strike
B <-35 #strike
pho <- -0.7

#Define functions
DOC_price_Hes<-function(S,K,B,r,sigma,t,pho){
nb_sim    <- 5000     # Number of Monte-Carlo simulations
nb_points <- 10^2      # Number of discretization points
payoff    <- c()       # Empty vector of simulated payoffs
a         <- 0.2
b         <- 0.2
c         <- 0.2
SIGMA <-rbind(c(1,pho),c(pho,1))
for(j in 1:nb_sim){
  x <- c(0)
  x <- S         # Initializing the process
  v <- sigma
  if (x<= B){
  payoff=0
  break
  }
    for(k in 2:nb_points){
      gauss<-mvrnorm(1,mu=c(0,0),Sigma=SIGMA)
      v[k] <- v[k-1]+(a-b*v[k-1])*t/nb_points+c*sqrt(v[k-1] * t/nb_points)* gauss[1]
      x[k] <- x[k-1] * (1 + r*t/nb_points + sqrt(v[k] * t/nb_points) *gauss[2]) #类似几何收益率的做法.
      tem=x[k] 
      
      if (x[k] <= B){
      tem=0
      break
      }
    }
  payoff[j] <- max(0,tem-K)
}

return(exp(-r*t)*mean(payoff))
}

#Get the price of DOC in Hestion model
Price_Hes=DOC_price_Hes(S,K,B,r,sigma,t,pho)
print(paste("Price of DOC in Hestion model: ", Price_Hes))

```



#####4. Visualize the simulation process
```{r}
nb_sim    <- 10    
nb_points <- 10^2 
a         <- 0.2
b         <- 0.2
c         <- 0.2
SIGMA     <- rbind(c(1,pho),c(pho,1))
x         <- rep(0,(nb_points*nb_sim))
dim(x)    <- c(nb_points,nb_sim)
v         <- rep(0,(nb_points*nb_sim))
dim(v)    <- c(nb_points,nb_sim)

#Generate 10 sample paths
for(j in 1:nb_sim){
  x[1,j]<- S         # Initializing the process
  v[1,j]<- sigma
  for(k in 2:nb_points){
    gauss<-mvrnorm(1,mu=c(0,0),Sigma=SIGMA)
    v[k,j] <- v[k-1,j]+(a-b*v[k-1])*t/nb_points+c*sqrt(v[k-1,j] * t/nb_points)* gauss[1]
    x[k,j] <- x[k-1,j] * (1 + r*t/nb_points + sqrt(v[k-1,j] * t/nb_points) *gauss[2]) #类似几何收益率的做法.
  }
}

#plot sample paths
sim <- data.frame(cbind(1:nb_points, x))
colnames(sim) <- c("Iteration", 1:nb_sim)
sim <- gather(sim, key = Simulation, value = Value, -Iteration)
sim$Simulation <- as.factor(sim$Simulation)
sim %>% ggplot(aes(x = Iteration, y = Value, color = Simulation)) + 
  geom_line() + theme(text = element_text(size = 13))+geom_hline(aes(yintercept=35),color ="red",linetype="dashed",size=1)+geom_hline(aes(yintercept=40),linetype="dashed",size=1)

#plot stock & vol together
Date <- (1:nb_points) / nb_points * t
x1<-x[,1]
Simul <- data.frame(Date, x1, "Stock")
colnames(Simul) <- c("Date", "Value", "Type")
Simul_vol <- data.frame(Date, v[,1], "Vol")
colnames(Simul_vol) <- c("Date", "Value", "Type")
Visu <- rbind(Simul, Simul_vol)
ggplot(Visu, aes(x = Date, y = Value)) + geom_line() + facet_grid(Type ~., scales = "free") +
      scale_x_continuous(minor_breaks = seq(0, 1, length = nb_points+1))

```

### 3.2.2 Robust Analysis for the price of DOC
Then, I will carry on the robust analysis in order to see how the Price of DOC will evolve with the price of the underlying stock, the barrier, the correlation coefficient, and the risk free rate respectively.

```{r}
#Price variation with initial stock prices
S_test=20:70
price<-c(0)
price=sapply(S_test,DOC_price_Hes,K=K, B=B, r=r, sigma=sigma, t=t,pho=pho)
price_test <- data.frame(S_test, price, "Price")
ggplot(price_test, aes(x = S_test, y = price)) + geom_line(color='red')

#Price variation with barrier
barrier_test=30+(1:50)/5
price<-c(0)
price=sapply(barrier_test,DOC_price_Hes,S=S,K=K,r=r, sigma=sigma, t=t,pho=pho)
price_test <- data.frame(barrier_test, price, "Price")
ggplot(price_test, aes(x = barrier_test, y = price)) + geom_line(color='red')

#Price variation with risk free rate
r_test=(1:50)/1000
price<-c(0)
price=sapply(r_test,DOC_price_Hes,S=S,K=K, B=B, sigma=sigma, t=t,pho=pho)
price_test <- data.frame(r_test, price, "Price")
ggplot(price_test, aes(x =r_test, y = price)) + geom_line(color='red')

#Price variation with correlation coefficient
pho_test=-(1:50)/50
price<-c(0)
price=sapply(pho_test,DOC_price_Hes,S=S,K=K, B=B, r=r,sigma=sigma, t=t)
price_test <- data.frame(pho_test, price, "Price")
ggplot(price_test, aes(x =pho_test, y = price)) + geom_line(color='red')

```


##3.3 Greeks
###3.3.1 Computing Greeks
The Greeks: Delta, Gamma, Vega, Theta, Pho) are also computed using the method mentioned in 2.3.1, and the results are as follows. h=0.001 are used here.
```{r}
#Definition of the function
DOC_price_Hes_G<-function(S,K,B,r,sigma,t,pho){
nb_sim    <- 100     # Number of Monte-Carlo simulations
nb_points <- 10      # Number of discretization points
payoff    <- c()       # Empty vector of simulated payoffs
a         <- 0.2
b         <- 0.2
c         <- 0.2
SIGMA <-rbind(c(1,pho),c(pho,1))
for(j in 1:nb_sim){
  x <- c(0)
  x <- S         # Initializing the process
  v <- sigma
  if (x<= B){
  payoff=0
  break
  }
    for(k in 2:nb_points){
      gauss<-mvrnorm(1,mu=c(0,0),Sigma=SIGMA)
      v[k] <- v[k-1]+(a-b*v[k-1])*t/nb_points+c*sqrt(v[k-1] * t/nb_points)* gauss[1]
      x[k] <- x[k-1] * (1 + r*t/nb_points + sqrt(v[k] * t/nb_points) *gauss[2]) #类似几何收益率的做法.
      tem=x[k] 
      
      if (x[k] <= B){
      tem=0
      break
      }
    }
  payoff[j] <- max(0,tem-K)
}

return(exp(-r*t)*mean(payoff))
}


Delta_price_Hes <- function(S,t){
               h<-0.1
          nb_sim<-500
          Delta <- c()  
          for(j in 1:nb_sim){
          Delta[j] <- DOC_price_Hes_G(S+h, K, B, r, sigma, t,pho)-DOC_price_Hes_G(S-h, K, B, r, sigma, t,pho)#center
}
  return(mean(Delta)/2/h)
}


Gamma_price_Hes <- function(S){
               h<-0.1
          nb_sim<-500
          Gamma <- c()  
          for(j in 1:nb_sim){
          Gamma[j] <- DOC_price_Hes_G(S+h, K, B, r, sigma, t,pho)+DOC_price_Hes_G(S-h, K, B, r, sigma, t,pho)-2*DOC_price_Hes_G(S-h, K, B, r, sigma, t,pho)#center
}
  return(mean(Gamma)/2/h^2)
}


Vega_price_Hes <- function(S){
               h<-0.001
          nb_sim<-500
          Vega <- c()  
          for(j in 1:nb_sim){
          Vega[j] <- DOC_price_Hes_G(S, K, B, r, sigma+h, t,pho)-DOC_price_Hes_G(S, K, B, r, sigma-h, t,pho)#center
}
  return(mean(Vega)/2/h)
}


Theta_price_Hes <- function(S){
               h<-0.001
          nb_sim<-500
          Theta <- c()  
          for(j in 1:nb_sim){
          Theta[j] <- DOC_price_Hes_G(S, K, B, r, sigma, t,pho)-DOC_price_Hes_G(S, K, B, r, sigma, t-h,pho)
}
  return(mean(Theta)/h)
}


Pho_price_Hes <- function(S){
             h<-0.0001
        nb_sim<-500
          Pho <- c()  
          for(j in 1:nb_sim){
          Pho[j] <- DOC_price_Hes_G(S, K, B, r+h, sigma, t,pho)-DOC_price_Hes_G(S, K, B, r-h, sigma, t,pho)#center
}
  return(mean(Pho)/2/h)
}


#Calculate Greeks
Delta_Hes=Delta_price_Hes(S,t)
Gamma_Hes=Gamma_price_Hes(S)
Vega_Hes=Vega_price_Hes(S)
Theta_Hes=Theta_price_Hes(S)
Pho_Hes=Pho_price_Hes(S)

#Output
print(paste("Delta: ",Delta_Hes))
print(paste("Gamma:",Gamma_Hes))
print(paste("Vega:",Vega_Hes))
print(paste("Theta:",Theta_Hes))
print(paste("Pho:",Pho_Hes))

```


###3.3.2 Robust Analysis for Delta
Here comes the robust analysis for Delta in four different time to maturity with h=0.001. 
```{r}
#Generate 4 curves
S_test=20:70
t_test=c(0.25,0.6,0.75,1)
Delta_test1=sapply(S_test,Delta_price_Hes,t=t_test[1])
Delta_test2=sapply(S_test,Delta_price_Hes,t=t_test[2])
Delta_test3=sapply(S_test,Delta_price_Hes,t=t_test[3])
Delta_test4=sapply(S_test,Delta_price_Hes,t=t_test[4])
Delta_test=cbind(Delta_test1,Delta_test2,Delta_test3,Delta_test4)

#Plot
test_greek <- data.frame(S_test, Delta_test)
colnames(test_greek) <- c("Stock_price", (1:4)/4)
test_greek<-gather(test_greek, key =Time_to_maturity, value = Value, -Stock_price)
test_greek$Time_to_maturity<- as.factor(test_greek$Time_to_maturity)
test_greek %>% ggplot(aes(x = Stock_price, y = Value, color = Time_to_maturity)) + 
  geom_line() + theme(text = element_text(size = 13))

```



If we take h=1 and T=1, we could get the following graph. 
```{r}
Delta_price_Hes <- function(S,t){
               h<-1
          nb_sim<-500
          Delta <- c()  
          for(j in 1:nb_sim){
          Delta[j] <- DOC_price_Hes_G(S+h, K, B, r, sigma, t,pho)-DOC_price_Hes_G(S-h, K, B, r, sigma, t,pho)#center
}
  return(mean(Delta)/2/h)
}

Delta_test=sapply(S_test,Delta_price_Hes,t=t) 
test_greek <- data.frame(S_test, Delta_test, "Delta")
ggplot(test_greek, aes(x = S_test, y = Delta_test)) + geom_line(color='red')

```